### 2018-07-17 Thesis/Parts/Abstract/Abstract.tex
Publishing note updates, all_latex added for Jenny
updates

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\fakeheader{Abstract}
Over the last few years, word embeddings have taken a dominant position in the Information Retrieval domain. Many studies have been done concerning the quality and application of word embeddings on general texts, such as the Wikipedia corpus and comments on review websites. Giving promising results, the word embeddings have been studied and improved over recent years. However, these studies have been focused on generic texts, which are not limited to the characteristics of in-domain texts such as rare domain-specific words or have been focussed on small sets of academic texts. This research focusses on the quality and application of word embeddings on domain-specific texts, concerning a large corpus of 1.391.543 scientific articles which have been published in 2017.
\end{document}

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
Over the last few years, word embeddings have taken a dominant position in the Information Retrieval domain. Many studies have been done concerning the quality and application of word embeddings on general texts, such as the Wikipedia corpus and comments on review websites. Giving promising results, the word embeddings have been studied and improved over recent years. However, these studies have been focused on generic texts, which are not limited to the characteristics of in-domain texts such as rare domain-specific words or have been focussed on small sets of academic texts. This research focusses on the quality and application of word embeddings on domain-specific texts, concerning a large corpus of 1.391.543 scientific articles which have been published in 2017. We validate the word embeddings using a categorization task to match articles to journals. We furthermore create a 2-dimensional visualization of the word embeddings on journal level, to visualize journal relatedness.\\
\vspace*{\fill}
\\
We aim to have this thesis published in the proceedings of the \textit{ECIR} (\url{http://ecir2019.org/}). \textit{Submission deadline: 9-10-2018}.
\end{document}

### 2018-07-17 Thesis/Parts/Data/Data.tex

The dataset we used for this research consists of articles published in 2017 which have been published in journals that have, in 2017, at least 150 publications. This results in a total dataset of $1.391.543$ articles from $3.759$ journals. Details about the corpus can be found in table~\ref{table:corpusSize}.

@@

\begin{tabular}{|l|c|c|c|}

@@

Total tokens & $186.962.354$ & $763.475$ & 134 \\

@@

\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their models show that, due to the additional parameters in the Pareto-III\footnote{The pareto distribution is a distribution in which recognizable by a long "tail" of low values and a quick and short ascend to the top values. More information: \url{https://en.wikipedia.org/wiki/Pareto_distribution}.}, the tail of the data fits better with the model than the Zipf\footnote{Distribution similar to the Pareto distribution, see: \url{https://en.wikipedia.org/wiki/Zipf\%27s_law}} model. This shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare, scientific texts and novels\cite{thurner2015understanding}.
The word occurrences in our corpus also follow the pattern of a Pareto distribution as described by~\citet{wiegand2018word}. Our distribution is visualized in Figure~\ref{figure:wordTokenOccurrence}, which displays the occurrences of the first 500 tokens of the corpus.

@@

\caption{Word and token occurrences. The word counts pattern resembles a Pareto distribution; the tokenization reduced the number of high-frequency words.}\label{figure:wordTokenOccurrence}

@@

The following steps have been applied to the words to create a tokenized set:

@@

For this research, we reused the word embeddings created by \citet{Truong2017Thesis}. These embeddings have a vector dimension of 300, which is an industry default. They have been trained on the entire Elsevier corpus (titles and abstracts respectively), not limited to the subset we used for this research. To create higher-level embeddings, we average all component embeddings. Thus, to create article embeddings, we take the average of all normalized word embeddings for that article. Journal embeddings are created by taking the average of all normalized article embeddings. We use the average to combine multiple embeddings into one because we determine the meaning of a larger text as the average meaning of all components. Since this meaning is represented as an embedding, we can average the embeddings to create the "average meaning". We have used multiple embedding optimizations for this research.

@@

The default embedding is created form the pre-trained word embeddings; no modifications have been applied to this set. We refer to this set in the figures as \textbf{embedding}.

@@

The TF-IDF weighted embedding set, referred to as TF-IDF embedding, are the default word embeddings weighted with a TF-IDF score per word.\\

@@

The articles embeddings are a normalized summation of each word vector multiplied by it TF-IDF value. Since we take a sum of all words, the Term Frequency is embedded as the raw count of each word. We will refer to this embedding in the figures as \textbf{tfidf\_embedding}.

@@

The 10K embedding set is generated similarity to the TF-IDF embedding, this version only uses the 10.000 most common tokens, reducing the number of tokens it uses. This set was created to see if the limitation to 10.000 tokens reduces the amount of noise, increasing the performance. We will refer to this embedding in the figures as the \textbf{10k\_embedding}.

@@

The 1K-6K TF-IDF embedding is the TF-IDF embedding limited to the top 6.000 most common words, without the top 1.000 most common words. The rationale for this is that common words will occur in many articles, creating noise, by cutting off the top 1.000 and cutting off everything below 6.000 we tried to reduce the noise by filtering common words. This cut results in a set of 5.000 tokens, which allows us to compare it to the 5K TF-IDF set. We will refer to this embedding in the figures as \textbf{1k\_6k\_embedding}. We refer to the \textbf{10k\_embedding}, \textbf{5k\_embedding} and the \textbf{1k\_6k\_embedding} as the \textbf{limited TF-IDF embeddings}, because these embedding use TF-IDF weighing, and have been limited in their vocabulary due to the cut-off's.

@@

To create the TF-IDF feature vectors, we used the TF-IDF model and a hasher from PySparks MlLib library. The TF-IDF feature vectors are created by hashing the tokens with the hasher, which has a set hash bucket size. These hashed values are passed on to the TF-IDF model, resulting in a feature vector which vector dimensions equal the number of hash buckets. To limit the computational and storage expenses and to reduce noise by rare words, we limit our vocabulary size.

@@

\caption{TF-IDF performance on title and abstract}\label{figure:tfidfPerformance}

@@

\caption{TF-IDF memory usage for title and abstract combined}\label{figure:tfidfMemory}

@@

Figure~\ref{figure:tfidfPerformance} shows the performance of title and abstract as median rank, and Figure~\ref{figure:tfidfMemory} shows the storage size in gigabyte, of the 1k/1K, 4K/4K, 7K/7K and 10K/10K TF-IDF configurations. This plot shows that, while the required storage size stagnates, the performance on title also quickly stagnates, together with the performance on the abstracts. The stagnation of the storage size is likely due to the fact that the terms which are added to the higher vocabulary sets occur less often than the other words (the cut-off is based on word occurrence, descending). Given this information, we have chosen to use the 10K/10K, 10K/5K and 5K/5K configurations to compare our embedding results to. Since these configurations can be kept in memory (storage size) while it is not likely that larger sets will give much improvement. The TF-IDF features are created on article level. We average the set of article feature vectors to create a journal feature vector, just as we did with the embeddings. 

@@

The research has been done in a python-databricks environment, which used Spark, a library that offers tools to work with big-data. \citet{armbrust2015spark} state that Spark SQL lets programmers leverage the benefits of relational processing and lets SQL users call complex analytic libraries in Spark. This allows for much tighter integration between relational and procedural processing. The paper further states that Spark SQL makes it significantly simpler and more efficient to write data pipelines that mix relational and procedural processing while offering substantial speedups over previous SQL-on-Spark engines.\\
We processed the TF-IDF sets and the embedding sets via the same pipeline, using their common vector properties. This ensures comparable results, the pipeline is set-up as follows:

@@

We split our initial set  $80\%$ - $20\%$. We use the $80\%$ set as the training set for the journal representations, and the $20\%$ set as the validation set for the journal representations. This split is based on a random number given to each article, ensuring that all set have the same (random) training and validation set.

@@

From our training set we create the journal embeddings, which are created for most sets\footnote{see paragraph datasets} by averaging the article embeddings or feature vectors.

@@

To categorize the articles, we calculate the distance between the title of the article, and the title of each journal from the validation set. We also do this for the abstract of the article and the abstract of each journal. During this process we keep track of:

@@

To calculate the distance between two vectors, cosine similarity is commonly used. We validated the quality of cosine similarity as a distance metric by comparing it to other similarity matrices available in the SciPy library\footnote{\url{https://www.scipy.org/}}, which we used to calculate the distances. We calculate the similarities based on the normalized embeddings, and compared the distance metrics based on the default embedding set. Table~\ref{table:distanceMetrics} shows the results of this validation.\\

@@

These results show high similarity between cosine-based metrics (Cosine \& Correlation) and euclidean based metrics (Euclidean, Seuclidean \& Sqeuclidean). This similarity is expected, since the cosine and euclidean distances should yield the same results on normalized sets. The results show that some enhancement on the euclidean algorithm result in slightly improved results, although not significant. Also the Cityblock metric yields results close to the Cosine metric, it has a slightly worse performance, which is also not significant. Because of this, we will use the cosine-similarity as the distance metric, which will make our results better comparable with other work.

@@

We define the positive \& negative metrics as follows:\\

@@

We used these metrics to calculate the Recall, Precision \& F1 as follows:\\

@@

$Recall = \dfrac{True Positive}{True Positive + False Negative}$\vspace{0.1in}\\
$Precision = \dfrac{True Positive}{True Positive + False Positive}$\vspace{0.1in}\\
$F1 = \dfrac{2 * Precision * Recall}{Precision + Recall}$

@@

@@

The median rank of a journal indicates the point where 50\% of all articles are to the left of a virtual line and 50\% are to the right of this line. The median rank therefore indicates the position of the "normal/typical"  article. This median rank differs from the average, since the median is not influenced by a single value. If the median is 20, it does not matter if an article rank is 25 of 250, it does matter however if an article rank is 19 or 21. For the average this is not the case, the average is influenced by single articles. Therefore, it does not represents the "normal/typical" article but it represents the exact average of all articles in the validation set, which is changed when an article rank is changed from 25 to 250.\\
We use the median rank to indicate at which rank the typical article would be ranked, based on its title and abstract. We do this by taking the median of the respective rank from each article. This gives us an indication of the behaviour of the articles in our validation set. We also calculate the average score for the title and abstract of all articles. This gives us four scores: \textbf{median title rank, median abstract rank, average title rank \& average abstract rank.}

@@

To further analyse the ranking results, we plot the rank distribution to get an indication of the ranking-landscape.

@@

To create a plot, we transformed the 300-dimensional journal vectors into 2-dimensional vectors using TSNE based on pca\footnote{Principal component analysis}. These 2-dimensional vectors, representing the x \& y coordinate, can then be drawn in a plot. To visualize the preservation of journal-relatedness while converting the 300 dimensions to 2 dimensions, we create groupings using k-means. These groupings are created on the 300-dimensional vectors and are visualized in the plot using colors. We use the k-means groupings due to the lack of subject-based grouping values in our dataset, for this research we used 8 groups.

>>>>

The dataset we used for this research consists of articles published in 2017 which have been published in journals that have (in 2017) at least 150 publications. This selection leads to a dataset of $1.391.543$ articles from $3.759$ journals. Details about the corpus can be found in table~\ref{table:corpusSize}.

@@

\begin{tabular}{|l|r|r|r|}

@@

Total tokens & $186.962.354$ & $763.475$ & 135 \\

@@

\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their work shows that, due to the additional parameters in the Pareto-III\footnote{The pareto distribution is a distribution in which recognizable by a long "tail" of low values and a quick and short ascend to the top values. More information: \url{https://en.wikipedia.org/wiki/Pareto_distribution}.}, the tail of the data fits better with the model than the Zipf\footnote{Distribution similar to the Pareto distribution, see: \url{https://en.wikipedia.org/wiki/Zipf\%27s_law}} model. The fit to the Pareto distribution shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare, scientific texts and novels\cite{thurner2015understanding}.
The word occurrences in our corpus also follow the pattern of a Pareto distribution as described by~\citet{wiegand2018word}. Our word and token distribution is visualized in Figure~\ref{figure:wordTokenOccurrence}, which displays the occurrences of the first 500 tokens of the corpus.

@@

\caption{Word and token occurrences. The word counts pattern resembles a Pareto distribution; the tokenization reduced the number of high-frequency words. The y-axis shows the number of occurances for each word, the x-axis shows the occurence rank of the word, meaning the position of the word in a list sorted on word occurance.}\label{figure:wordTokenOccurrence}

@@

The following steps have been applied to the words from the text to create a tokenized set:

@@

For this research, we reused the word embeddings created by \citet{Truong2017Thesis}. These embeddings have a vector dimension of 300, which is an industry default. They have been trained on the entire Elsevier corpus (titles and abstracts respectively), not limited to the subset we used for this research. To create higher-level embeddings, we average all component embeddings. Thus, to create article embeddings (higher level), we take the average of all normalized word embeddings (component) for that article. Journal embeddings are created by taking the average of all normalized article embeddings. We use the average to combine multiple embeddings into one because we determine the meaning of a larger text as the "average meaning" of all components. Since this meaning is represented as an embedding, we can average the embeddings to create the average meaning. We have used multiple embedding optimizations for this research.

@@

The default embedding is created from the pre-trained word embeddings; no modifications have been applied to this set. We refer to this set in the figures as \textbf{embedding}.

@@

The TF-IDF weighted embedding sets, referred to as TF-IDF embedding, are the default word embeddings weighted with a TF-IDF score per word.\\

@@

The article embeddings are a normalized summation of each word vector multiplied by it TF-IDF value. Since we take a sum of all words, the Term Frequency is embedded as the raw count of each word. We will refer to this embedding in the figures as \textbf{tfidf\_embedding}.

@@

The 10K embedding set is generated similarity to the TF-IDF embedding, this version only uses the 10.000 most common tokens, reducing the number of tokens it uses. This set was created to see if the limitation to 10.000 tokens reduces the amount of noise\footnote{Noise in this context are rare word}, increasing the performance. We will refer to this embedding in the figures as the \textbf{10k\_embedding}.

@@

The 1K-6K TF-IDF embedding is the TF-IDF embedding limited to the top 6.000 most common words, without the top 1.000 most common words. The rationale for this cutting is that common words will occur in many articles, creating noise, by cutting off the top 1.000 and cutting off everything below 6.000 we tried to reduce the noise by filtering common words. This cut results in a set of 5.000 tokens, which allows us to compare it to the 5K TF-IDF set. We will refer to this embedding in the figures as \textbf{1k\_6k\_embedding}. We refer to the \textbf{10k\_embedding}, \textbf{5k\_embedding} and the \textbf{1k\_6k\_embedding} as the \textbf{limited TF-IDF embeddings}, because these embedding use TF-IDF weighing, and have been limited in their vocabulary due to the cut-off's.

@@

To create the TF-IDF feature vectors, we used the TF-IDF model and a hasher from PySparks MlLib library\footnote{\url{http://spark.apache.org/docs/2.0.0/api/python/pyspark.mllib.html}.}. The TF-IDF feature vectors are created by hashing the tokens with the hasher, which has a set hash bucket size. These hashed values are passed on to the TF-IDF model, resulting in a feature vector. The vector dimensions of the feature vector equal the number of hash buckets. To limit the computational and storage expenses and to reduce noise by rare words, we limit our vocabulary size.

@@

\caption{TF-IDF performance on title and abstract.}\label{figure:tfidfPerformance}

@@

\caption{TF-IDF memory usage for title and abstract combined.}\footnote{We combined the memory usage of the titles and abstracts since the titles and abstract are created together and share their feature vector size.}\label{figure:tfidfMemory}

@@

Figure~\ref{figure:tfidfPerformance} shows the performance of TF-IDF on title and abstract as median rank, and Figure~\ref{figure:tfidfMemory} shows the storage size in gigabyte, of the 1k/1K, 4K/4K, 7K/7K and 10K/10K TF-IDF configurations. This plot shows that, while the required storage size stagnates, the performance on title also quickly stagnates; the performance of the abstracts show similar behaviour. The stagnation of the storage size is likely due to the fact that the terms which are added to the larger vocabulary sets occur less often than the other words (the cut-off is based on word occurrence, descending). Given this information, we have chosen to use the 10K/10K, 10K/5K and 5K/5K configurations to compare our embedding results to because these configurations can be kept in memory (storage size) while it is not likely that larger sets will give much improvement. The TF-IDF features are created on article level. We average the set of article feature vectors to create a journal feature vector, just as we did with the embeddings. 

@@

The research has been done in a python-databricks environment, which used Spark SQL, a library that offers tools to work with big-data. \citet{armbrust2015spark} state that Spark SQL lets programmers leverage the benefits of relational processing and lets SQL users call complex analytic libraries in Spark. This allows for much tighter integration of relational and procedural processing. The paper further states that Spark SQL makes it significantly simpler and more efficient to write data pipelines that mix relational and procedural processing while offering substantial speedups over previous SQL-on-Spark engines.\\
\subheader{Data processing}
We processed the TF-IDF sets and the embedding sets \footnote{See Chapter 3.2 Datasets.} via the same pipeline, using their common vector properties. Using the same pipeline ensures comparable results. The pipeline is set-up as follows:

@@

We split our initial set  $80\%$ - $20\%$. We use the $80\%$ set as the training set for the journal representations, and the $20\%$ set as the validation set for the journal representations. This split is based on a random number given to each article, ensuring that all datasets (i.e. tfidf\_10k\_10k, embedding) have the same (random) training and validation set.

@@

From our training set we create the journal embeddings, which are created for most sets\footnote{See chapter 3.2 Datasets.} by averaging the article embeddings or feature vectors.

@@

To categorize the articles, we calculate the distance between the title of the article, and the title of each journal from the validation set. We also do this for the abstract of the article and the abstract of each journal. During this process we record of:

@@

To calculate the distance between two vectors, cosine similarity is commonly used. We validated the quality of cosine similarity as a distance metric by comparing it to other similarity metrics available in the SciPy library\footnote{\url{https://www.scipy.org/}}. We calculate the similarities based on the normalized embeddings, and compared the distance metrics based on the default embedding set. Table~\ref{table:distanceMetrics} shows the results of this validation.\\

@@

These results show high similarity between cosine-based metrics (Cosine \& Correlation) and euclidean based metrics (Euclidean, Seuclidean \& Sqeuclidean). This similarity is expected, since the cosine and euclidean distances should yield the same results on normalized sets\footnote{See: \url{https://en.wikipedia.org/wiki/Cosine_similarity\#Properties}.}. The results show that some enhancement on the euclidean algorithm (Seuclidean) result in slightly improved results, although not significant. Also the Cityblock metric yields results close to the Cosine metric, only having a slightly worse performance, which is also not significant. Because of this, we will use the cosine-similarity as the distance metric, which will make our results easier to compare with other work, since the cosine-similarity is a commonly used similarity measure.

@@

For our matching algorithm, we classify the results as follows:

@@

With these definitions, we calculate the Recall, Precision \& F1 as follows:\\

@@

$Recall = \dfrac{|True Positive|}{|True Positive| + |False Negative|}$\vspace{0.1in}\\
$Precision = \dfrac{|True Positive|}{|True Positive| + |False Positive|}$\vspace{0.1in}\\
$F1 = \dfrac{2 * Precision * Recall}{Precision + Recall|}$

@@

here we denote the amount of articles in a class as $|class|$.

@@

A dataset rank is calculated using the ranks of all article in that dataset. The median rank of a dataset indicates the point where 50\% of all articles are to the left of a virtual line and 50\% are to the right of this line. The median rank therefore indicates the position of the "normal/typical" article. This median rank differs from the average, since the median is not influenced by a single value. If the median is 20, it does not matter if an article rank is 25 or 250, it does matter however if an article rank is 19 or 21. For the average rank this is not the case, the average rank is influenced by single articles. Therefore, it does not represents the "normal/typical" article but it represents the exact average rank of all articles in the validation set, which is changed when an article rank is changed from 25 to 250.\\
We use the median rank to indicate at which rank the typical article would be ranked, based on its title and abstract. We also calculate the average score for the title and abstract of all articles. This gives us four scores: \textbf{median title rank, median abstract rank, average title rank \& average abstract rank.}

@@

To further analyse the ranking results, we plot the article rank distribution to get an indication of the ranking-landscape.

@@

To create a plot, we transformed the 300-dimensional journal vectors into 2-dimensional vectors using TSNE based on PCA\footnote{Principal component analysis}. These 2-dimensional vectors, representing the x \& y coordinate, can then be drawn in a plot. To visualize the preservation of journal-relatedness while converting the 300 dimensions to 2 dimensions, we create groupings using k-means. These groupings are created on the 300-dimensional vectors and are visualized in the plot using colors. We use the k-means groupings due to the lack of subject-based grouping values in our dataset, for this research we used 2, 5, 8, 10, 20, 25, 30, 35 and 40 as the amount of groups. We included only the plots of 8 groups, because we find this plot the most informative. All group amounts can be seen in the digital version of the plots.
